Experience shapes individual foraging specialization and success in a virtual predator-prey system

### 2023-10-17 manuscript/Chapter2-manuscript.Rmd
updt 1st paragraph of intro for 3rd submission

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing on different foraging tactics, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Despite clear evidence that individual specialization is important in predator-prey interactions, its underlying mechanisms remain surprisingly unclear. One hypothesis suggests that reciprocal interactions between predators and prey could drive individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022]. For instance, predators often hunt using techniques that are fine-tuned to the type of prey that they encounter [@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their refinement is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Thus, the capacity of predators to effectively use hunting tactics should be contingent on periods of extensive practice (i.e. experience), which may result in different foraging specializations.

@@

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging strategies [@Phillips.etal2017]. Some studies have found greater success for individuals specializing on a foraging tactic [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others have found that individuals using flexible foraging tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021]. This discrepancy may in part be due to the limited information we have on the ecological contexts favoring individual specialization over flexibility. However, many studies also show that both type of foragers can attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018]. It is hypothesized that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit from adjusting to fluctuating resources by reducing the consequences of uncertainty [@dall2010]. While it is clear that predators can match their hunting behaviour and locomotion to their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022], the important question remaining is how. Given the dynamic nature of predator-prey interactions, it is likely that ontogenic behavioural changes in predators play an important role, because complex hunting strategies require practice to be effective [@Kelley.Magurran2011]. Uncovering the role of experience in shaping direct predator-prey interactions would help to predict the behavioural decisions and the success of predators when they are hunting.

>>>>

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals using different foraging tactics to hunt their prey, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Theory suggests that foraging tactics are learned through repeated interactions with prey, and the capacity of predators to effectively use them should be contingent on periods of extensive practice, i.e., experience [@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022;@Tinker.etal2009;@Dukas2019]. Empirical studies show that predators often hunt using techniques that are fine-tuned to the type of prey that they encounter, and their refinement is crucial for young predators to reach adulthood and survive [@Estes.etal2003;@Woo.etal2008;@Phillips.etal2017;@Heithaus.etal2018;@Courbin.etal2018]. However, the underlying mechanisms shaping these processes remain unclear, which hinders our ability to predict the behavioural decisions and the success of predators when they are hunting.

@@

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging strategies [@Phillips.etal2017]. Some studies have found greater success for individuals specializing on a foraging tactic [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others have found that individuals using flexible foraging tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021]. This discrepancy may in part be due to the limited information we have on the ecological contexts favoring individual specialization over flexibility. However, many studies also show that both type of foragers can attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018]. It is hypothesized that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit from adjusting to fluctuating resources by reducing the consequences of uncertainty [@dall2010]. While it is clear that predators can match their hunting behaviour and locomotion to their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022], the important question remaining is how. Given the dynamic nature of predator-prey interactions, it is likely that ontogenic behavioural changes in predators play an important role, because complex hunting strategies require practice to be effective [@Kelley.Magurran2011].
